Gravitaitonal Wave Event Progenitor Synthesis

### 2022-05-26 src/tex/ms.tex
Will further edits intro for clarity and adds some citations.

Traditionally, population synthesis studies simulate merging DCO populations
with a Monte-Carlo approach.  Using theoretically motivated distributions of
initial parameters like age, metallicity, component mass, orbital separation,
and eccentricity, initial populations are evolved with fixed astrophysical
assumptions, or population synthesis parameter settings, to produce a synthetic
catalog of DCO mergers which can be compared to parameterized models derived
from observations. This approach relies on several assumptions, both in the
simulations and model parameterizations fitted to the GW data, which are
unlikely to be fully correct.  For example, there is no reason to suppose that
the entire population of DCOs should evolve with the same fixed hyperparameters;
there is no a priori reason the fitted parameterized model should capture the
relevant features of the observed DCO distributions to correspond to the binary
physics of interest; etc. A better approach would be to compare progenitor
parameters and hyperparameter settings directly to DCO observations, treating
the population synthesis model as a mapping from progenitor to merger
parameters; merger parameters can then be mapped into observations using a
waveform family, and parameter inference proceeds in the usual way
\citep[e.g.][]{Veitch2015}.  A similar methodology, though with \emph{fixed}
hyperparameter settings, was advanced in \citet{Andrews2018,Andrews2021}, as we
discuss in more detail below.

%One of the most common skepticisms of population synthesis studies is that they employ simulation sets which often
%cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
%(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
%On top of this, these values are often not calibrated to data.
Attempting to approach such an improved inference procedure for evolutionary
hyperparameters using traditional en mass Monte Carlo simulations of entire
populations of DCO mergers is computationally infeasible. Even with recent
developments in using emulators to speed up the simulation process
\citep[e.g.][]{Wong2021}, training the emulators still requires a significant
amount of computation to cover a wide range of uncertainty in the hyperparemeter
space. As an example, to emulate a model with $10$ parameters, the simplest way
to construct a training set of simulations is to run simulations on a grid which
varies combinations of uncertain physics. For $10$ uncertain physical processes,
if we consider $2$ variations for each physical process this would still require
$1024$ simulations. Because population synthesis simulations require hundreds to
thousands of cpu hours, training an emulator which spans a high dimensional
uncertainty space remains an impossibility at present.  Thus inference of
hyperpameter settings must proceed \emph{dynamically}, generating trial DCO
systems to match a particular observation one- or few-at-a-time.

>>>>

Traditionally, population synthesis studies simulate merging DCO populations
with a Monte-Carlo approach.  Using theoretically motivated distributions of
initial parameters like age, metallicity, component mass, orbital separation,
and eccentricity, initial populations are evolved with fixed astrophysical
assumptions, or population synthesis hyperparameter settings, to produce a
synthetic catalog of DCO mergers which can be compared to parameterized models
derived from observations. This approach relies on several assumptions, both in
the simulations and model parameterizations fitted to the GW data, which are
unlikely to be fully correct.  For example, there is no reason to suppose that
the entire population of DCOs should evolve with the same fixed hyperparameters;
there is no a priori reason the fitted parameterized model should capture the
relevant features of the observed DCO distributions to correspond to the binary
physics of interest; etc. A better approach would be to compare progenitor
parameters and hyperparameter settings directly to DCO observations, treating
the population synthesis model as a mapping from progenitor to merger
parameters; merger parameters can then be mapped into observations using a
gravitational waveform family, and parameter inference proceeds in the usual
way, propagating information back up the mapping chain
\citep[e.g.][]{Veitch2015}.  A similar methodology, though with \emph{fixed}
hyperparameter settings, was advanced in \citet{Andrews2018,Andrews2021}, as we
discuss in more detail below.

%One of the most common skepticisms of population synthesis studies is that they employ simulation sets which often
%cover a small subdomain of the entire parameter space of uncertain physics, or only vary one uncertain process per simulation
%(e.g. common envelope efficiency) while holding all others fixed to some ``default" values.
%On top of this, these values are often not calibrated to data.
Attempting to approach such an improved inference procedure for evolutionary
hyperparameters using traditional en mass Monte Carlo simulations of entire
populations of DCO mergers is computationally infeasible. Even with recent
developments in using emulators to speed up the simulation process
\citep[e.g.][]{Wong2021}, training the emulators still requires a significant
amount of computation to cover a wide range of uncertainty in the hyperparemeter
space. As an example, to emulate a model with $10$ parameters, the simplest way
to construct a training set of simulations is to run simulations on a grid which
varies combinations of uncertain physics. For $10$ uncertain physical processes,
if we consider $2$ variations for each physical process this would still require
$1024$ simulations. Because population synthesis simulations require hundreds to
thousands of cpu hours, training an emulator which spans a high dimensional
uncertainty space remains an impossibility at present.  Thus inference of
hyperpameter settings must proceed \emph{dynamically}, generating trial DCO
systems via population synthesis recipes to match a particular observation one-
or few-at-a-time while simultaneously adjusting the hyperparameters. (Eventually
it may be even be possible to use more physically-motivated modeling of binary
physics \citep[e.g.][]{Gallego-Garcia2021} in such a procedure.)
